---
title: "Probability Simulation with R"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Question 1:**

In an office, there are 10 sales person,four women and six men. Three women and two men will be chosen at random. Use simulations in R to approximate the probability.

```{r 1,, echo=TRUE}
nloop=10000

sample1  = c(rep("W",4),rep("M",6))
sample1
count = 0

for (iloop in 1:nloop){
  MW = sample(sample1,5, replace=F)
  if  (sum(length(which("M"==MW)))==2){count =count+1}
}
count/nloop

#Mathematical Solution
Com = function(n, x) {
  factorial(n) / factorial(n-x) / factorial(x)
}

(Com(4,3) * Com(6,2))/ Com(10,5)

```

**Question 2:**

There are 7 balls in a urn, they are identical except color, 3 blue and 4 yellow. What is the probability of choosing 2 balls in a different color without replacement ? Write R code to approximate the probability. 


```{r 2, echo=TRUE}

nloop=10000
bag = c(rep(1,3),rep(0,4))
#Let's assign the colors to numbers: Blue=1 and Yellow=0
count=0

for (iloop in 1:nloop){
  tot = sample(bag,2, replace=F)
  if (sum(tot)==1 ){count =count+1}
}
count/nloop

#Mathematical Solution
Com(2,1)*(3/7)*(4/6)

```

**Question 3:**

There are 10 balls in each urn.  5 blue and 5 yellow ball in the 1st urn, and 4 blue, and 6 yellow in the
2nd urn. Sandy picks 1 ball from each urn at random. What is the probability of choosing the balls in the same color?
Write R code to approximate the probability.

```{r 3,, echo=TRUE}

# Let's code Blue= 1, Yellow= 0 
nloop=10000
sam1 = c(rep(1,5),rep(0,5))
sam2 = c(rep(1,4),rep(0,6))
count=0

for (iloop in 1:nloop){
  urn1 = sample(sam1,1, replace=F)
  urn2 = sample(sam2,1, replace=F)
  urn  = urn1 + urn2
  
  if (sum(urn)==2 | sum(urn==0)){count =count+1}
  
}
count/nloop

#Mathematical Solution
(5/10)*(4/10)+(5/10)*(6/10)

```

**Question 4:**
Suppose you roll a fair six-sided die six times. Write R code to approximate the
   probability that each number comes up exactly one time.

```{r 4, echo=TRUE}

nloop=10000
sam1 = rep(1:6,1)
count=0

for (iloop in 1:nloop){
  roll = sample(sam1,6, replace=T)
  sroll = sort(roll)
  
  if (all(sroll == c(1,2,3,4,5,6)) ){count =count+1}
  
}
count/nloop

#Mathematical Solution
(6/6)*(5/6)*(4/6)*(3/6)*(2/6)*(1/6)

```

**Question 5:**
What is the probability of a large straight when five dice are rolled? This is defined (in the game     Yahtzee) as the numbers 1,2,3,4,5 or 2,3,4,5,6. Write R code to approximate this probability.


```{r 5,, echo=TRUE}

nloop=10000
sam1 = rep(1:6,1)
count=0

for (iloop in 1:nloop){
  roll = sample(sam1,5, replace=T)
  sroll = sort(roll)
  
  if (all(sroll == c(1,2,3,4,5)) | all(sroll == c(2,3,4,5,6))){count =count+1}
  
}
count/nloop

#Mathematical Solution
2*(5/6)*(4/6)*(3/6)*(2/6)*(1/6)

```


**Question 6:**
 For a recent year, 51% of the families in the United States had no children under the age of 18; 
 20% had one child; 19% had two children; 7% had three children; and 3% had four or more children.  
 If three family is selected at random, find the probability that total number of children these families
 have less then 7. Write R code to determine the probability.
 Use simulations in R to approximate probability.


```{r 6, echo=TRUE}

nloop = 10000
count = 0

for (iloop in 1:nloop){
  child = sample(0:4,3,replace = T, prob = c(.51,.2,.19,.07,.03))
  if (sum(child)<7){count=count+1}
  
}
count/nloop

#Mathematical Solution
# We need to find all different possibilities and add them up. (000,001,002,...)

```


**Question 7:**
   Four children from a neighborhood in Smalltown attend a holiday party. There are
   40 children total at the party, and 20 prizes are to be distributed at random to
   the children. That is, each prize is equally likely to go to each of the 40 children,
   regardless of whether a child already has prizes.
   
  (a) Write R code to determine the probability that each of the four children gets
      exactly one prize.
   
 
```{r 7a,, echo=TRUE}

#a.

nloop =10000
count=0

for(iloop in 1:nloop){
  prize=sample(1:40,20,replace=T)
  s_prize = sort(prize)
  
  if (all(s_prize[1:4] == c(1,2,3,4)) & s_prize[5]!=4){count= count+1}
}
count/nloop

#Mathematical Solution
(20*19*18*17*(36)^16)/40^20

```


      
  (b) Write R code to determine the probability that each of the four children gets at
      least one prize.
      
```{r 7b, echo=TRUE}

nloop =10000
count=0

for(iloop in 1:nloop){
  price=sample(1:40,20,replace=T)
  
  if (all(c(1,2,3,4) %in% price)){count= count+1}
}
count/nloop

#Mathematical Solution

(20*19*18*17*(40)^16)/40^20
```

 (b) Write a R code to determine the probability that 50 different children gets a prize.
 
```{r 7b,, echo=TRUE}

#c. UNIQUE and  LENGTH

nloop =10000
count=0

for(iloop in 1:nloop){
  prize=sample(1:40,20,replace=T)
  u_prize =length(unique(prize))
 
  if (u_prize== 20){count= count+1}
}
count/nloop

#Mathematical Solution

(factorial(20)*Com(40,20))/40^20

```


**Question 8:**
 Two factories supply light bulbs to the market. Bulbs from factory X work  for over 6000 hours in 
 98% of cases, whereas bulbs from factory Y work for over 6000 hours  in 93% of cases. It is known that   
 factory X supplies 65% of the total bulbs available in the  market.
 Given that a light bulb works for more than 6000 hours, what is the probability that it
 was supplied by factory Y ?
 Use simulations in R to approximate probability.

```{r 8, echo=TRUE}


nloop   =10000
countx  =0
county  =0
countxy =0
for (iloop in 1:nloop){
  xy= sample(c("x","y"),1, prob=c(0.65,0.35))
  if (xy=="x"){ 
    x = sample(0:1,1, prob =c(0.02,0.98))
    if(x=="1"){countx =countx+1}
  }
  if (xy=="y"){
    y = sample(0:1,1, prob =c(0.07,0.93))
    if(y=="1"){county = county+1}
  }
  
}

county/(countx+county)

#Mathematical Solution
(93*35)/(98*65+93*35)

```

**Question 9:**
     
   A drawer contains 10 pairs of socks, where each pair is a different color. Sam draws
   four socks at random from the drawer (without replacement). Write R code to
   determine the probability that there is no pair among the four socks.


```{r 9,, echo=TRUE}

nloop =10000
count =0
socks = c(rep(1:10,each=2))

for (iloop in 1:nloop){
  draw_soc = sample(socks,4, replace=F)
  sock = unique(draw_soc)
  if(length(sock)==4){count= count+1}
  
}
count/nloop

#Mathmatical Solution:
(20/20)*(18/19)*(16/18)*(14/17)

```

**Question 10:**

  There are 4 girl and 5 boy in the classroom. What is the probability of having girls stand next to
  each other  in the line? Use simulations in R to approximate probability.

```{r 10, echo=TRUE}

nloop =10000
count =0
tot = c(1:9)

for (iloop  in (1:nloop-3)){
  samp = sample(tot,9, replace=F)
  for (i in  1:9){
    if(all(samp[i]   %in% c(1:4) & 
           samp[i+1] %in% c(1:4) & 
           samp[i+2] %in% c(1:4) &
           samp[i+3] %in% c(1:4))) {count =count+1}
  }
}
count/nloop

#Mathematical Solution

factorial(6)*factorial(4)/factorial(9)

```

**Question 11:**

  In a classroom, there are total 30 students and 8 of them have glasses, and half of the students 
  who wear and not wear glasses habe brown eyes.
  What is the probability of choosing a student who has a broen eye or wear glasses?


```{r 11,, echo=TRUE}

nloop=10000
count =0
class = c(rep(1,19),rep(0,11))

for (iloop in 1:nloop){
  bg= sample(class,1, replace=F)
  if(bg==1){count=count+1}
}

count/nloop

#Mathematical solution
19/30

```

**Question 12:**

  The probablity of a hunter to hit the targer is 3/4. If he makes 4 shoot, 
  what is the probability of hitting the target only on shoot 2 and shoot 4?

```{r 12, echo=TRUE}

nloop =10000
count =0

for (iloop in 1:nloop){
  hit1 = sample(0:1,1,replace= F,prob=c(.25,.75))
  hit2 = sample(0:1,1,replace= F,prob=c(.25,.75))
  hit3 = sample(0:1,1,replace= F,prob=c(.25,.75))
  hit4 = sample(0:1,1,replace= F,prob=c(.25,.75))
  if((sum(hit1+hit3)==0) & (sum(hit2+hit4)==2)){count=count+1}
}
count/nloop

#Mathematical Solution
(1/4)*(3/4)*(1/4)*(3/4)

```

**Question 13:**

  Two evenly matched baseball teams play each other for a series of seven games. Estimate the probability   that Team I will win the series by winning at least four games from Team 2. 
  Use simulations in R to approximate wining probability.


```{r 13,, echo=TRUE}

nloop = 10000
count =0


for (iloop in 1: nloop){
  win = sample(0:1,7,replace=T)
  if (sum(win)>=4){count =count+1}
}
count/nloop

#Mathematical Solution
com = function(n, x) {
  factorial(n) / factorial(n-x) / factorial(x)
}


com(7,4)*(1/2)^7+com(7,5)*(1/2)^7+com(7,6)*(1/2)^7+com(7,7)*(1/2)^7




```

**Question 14:**

   Five team joins a soccer tournamnet. The wining posibilities as follows: 
   Team A & Team B =60%,  Team A & Team C =40%, Team A & Team D =80%,   Team A & Team E =30%,
   Team B & Team C =30%,  Team C & Team D =70%, Team C & Team E =40%,
   Team C & Team D =50%,  Team C & Team E =80%, 
   Team D & Team E =90%,
   
   Use simulations in R to approximate the winning posibilities for each team.

```{r 14, echo=TRUE}
nloop =100
countA =0; countB =0; countC =0; countD =0; countE =0
for (i in 1:nloop){
  gameAB= sample(c(1,2),1, prob=c(0.6,0.4))
  gameAC= sample(c(1,3),1, prob=c(0.4,0.6))
  gameAD= sample(c(1,4),1, prob=c(0.8,0.2))
  gameAE= sample(c(1,5),1, prob=c(0.3,0.7))
  gameBC= sample(c(2,3),1, prob=c(0.3,0.7))
  gameBD= sample(c(2,4),1, prob=c(0.7,0.3))
  gameBE= sample(c(2,5),1, prob=c(0.4,0.6))
  gameCD= sample(c(3,4),1, prob=c(0.5,0.5))
  gameCE= sample(c(3,5),1, prob=c(0.8,0.2))
  gameDE= sample(c(4,5),1, prob=c(0.9,0.1))
  games =c(gameAB,gameAC,gameAD,gameAE,gameBC,gameBD,gameBE,gameCD,gameCE,gameDE)
  
  countA = countA + length(which(1==games))
  countB = countB + length(which(2==games))
  countC = countC + length(which(3==games))
  countD = countD + length(which(4==games))
  countE = countE + length(which(5==games))
  
}

print(paste("Team A :", countA,"// Team B :",countB, "// Team C :",countC,"// Team D :",countD,"// Team E :",countE))


```



**Question 15:**

The experiment is to roll four fair, six-sided dice. Given that all dice show numbers
that are three or greater, what is the probability that all the dice show either odd numbers or even numbers(3,5,7,5 or 4,6,4,6) ? Use simulations in R to compute this conditional probability.


```{r 15,, echo=TRUE}

nloop=10000; countA=0; countAB=0
for(iloop in 1:nloop){
  dice=sample(1:6,4,replace=TRUE)
  eventA =  all(dice>=3)
  eventB =  (all(dice%%2==1) | all(dice%%2==0))
  if(eventA){countA=countA+1}
  if(eventA&eventB){countAB=countAB+1}
}
countAB/countA

# Mathematical Solution

# Choosing odd numbers   =2^4
# Chosing even numbers  =2^4
# All possibilites that are greater or equal to 3 = (4)^4 =256
(2^4+2^4)/4^4


```



**Question 16:**

A quiz consists of ten multiple choice questions with choices (a) and (b)for
each. If an unprepared student marks answers at random, what are the probabilities
of the following events:

Use simulations in R to approximate probabilities.

(a) exactly one correct answer


```{r 16a, echo=TRUE}

nloop =10000
count =0

for (iloop in 1:nloop){
  quiz = sample(0:1,10,replace = T)
  #Let's choose 1's are correct and 0's are wrong answer
  if (sum(quiz)==1){count= count+1}
}

count/nloop

#Mathematical Solution
Com = function(n, x) {
  factorial(n) / factorial(n-x) / factorial(x)
}

(Com(10,1) )/ 2^10
```

(b) seven or more answers are correct

```{r 16b, echo=TRUE}

nloop =10000
count =0

for (iloop in 1:nloop){
  quiz = sample(0:1,10,replace = T)
  #Let's choose 1's are correct and 0's are wrong answer
  if (sum(quiz)>=7){count= count+1}
}

count/nloop

#Mathematical Solution
Com = function(n, x) {
  factorial(n) / factorial(n-x) / factorial(x)
}


(Com(10,7) + Com(10,8) + Com(10,9) + Com(10,10))/ 2^10
```


**Question 17:**

A quiz consists of five multiple choice questions with choices (a), (b) and (c)for
each. If an unprepared student marks answers at random, what are the probabilities
of the following events:

   Use simulations in R to approximate probabilities.

  (a) exactly one correct answer
 

```{r 17a, echo=TRUE}

nloop =10000
count =0

for (iloop in 1:nloop){
  quiz = sample(c("a",'b',"c"),5,replace = T)
  
  #Let's choose a's are correct and b's and c's are wrong answer
  if (quiz[1]=="a"){count= count+1}
}

count/nloop

#Mathematical Solution
Com = function(n, x) {
  factorial(n) / factorial(n-x) / factorial(x)
}


(Com(5,1)*(1/3)*(2/3)^4)


```

 (b) three or more answers are correct

```{r 17b, echo=TRUE}

nloop =10000
count =0

for (iloop in 1:nloop){
  quiz = sample(c("a",'b',"c"),5,replace = T)
  
  #Let's choose a's are correct and b's and c's are wrong answer
  if ((sum(sapply(quiz=="a",unique)))>=3){count= count+1}
 
}

count/nloop

#Mathematical Solution
Com = function(n, x) {
  factorial(n) / factorial(n-x) / factorial(x)
}


(Com(5,3)*(1/3)^3*(2/3)^2 + Com(5,4)*(1/3)^4*(2/3) + Com(5,5)*(1/3)^5)


```

**Question 18:**

Simulate Birthday paradox with R.

The paradox of birthdays is a mathematical problem put forward by Von Mises, who looks for the value N in the problem: In a group of N people there is 50% chance that at least 2 people in the group share the same birthday (day + month). The answer is N = 23, which is quite counterintuitive, hence the paradox.

```{r 20, echo=TRUE}

# Birthday Function
birthday <- function(n){
  f=1
  for(i in 1:n){
    f <- f* (365-i+1)/365
  }
  print(paste("Probability of",i,"people  :",1-f))
}

# Birthday Paradox Simulation
nloop=50
birthstat=1:nloop
for (i in 1:nloop){
  birthstat[i]=birthday(i)
}


# x=1:50
# plot(x,birthstat, main= "Birthday Paradox", xlab = "Number of people", ylab="Probability")
# abline(h=0.5, xlim =c(0,50),col="red")


```


